Expos\'{e} Master Thesis\\ Checking Assertions for Smart Contracts

### 2020-09-08 meeting-notes/20200812-expose.tex
update

@@

\section{Further Thoughts}
\label{sec:further-thoughts}

There may be an alternative way of implementing this scheme without modifying the
protocol, but instead adding instructions to the virtual machine. One instruction would
generate a random number (EVM uses 512 bit unsigned integer IIRC), the other would
raise an overriding failure. Ideally, once a counterexample has been found, this
counterexample is broadcast to the network where it can be verified by further nodes.
But this broadcast and the reaction to this broadcast requires an update of the
protocol: all receivers of the broadcast will check the counterexample and revert the
transaction. This probably means that a transaction using this feature has to wait for
a couple of cycles in the transaction pool until the node feels safe to include the
transaction in a block.

\begin{changethis}
  Tezos specific remarks: here the number of validators is really small compared to
  Bitcoin or Ethereum
  \begin{itemize}
  \item maybe we need to propose and validate the same block multiple times before it
    gets accepted. E.g, a transaction containing this feature is only accepted if it is
    revalidated $n$ times.
  \item Extra validators only for this kind of transaction.
  \end{itemize}
\end{changethis}
This approach poses the danger of a DOS attack. Someone could post a false
counterexample, which does not really lead to a failure. If everyone switches to check
this proposed counterexample instead of generating a random number, then the real
problem may remain hidden. So, there has to be a component in the protocol that
penalizes someone who broadcasts a false counterexample. Moreover, not all validators
should switch to recheck a proposed counterexample; or they could check it in addition
to their own random check.

The interesting assertion formulas are those with prenex universal
quantification. Negation turns them into prenex existential quantification. But it
would be ineffective to generate a random integer and then check that it falls in the
interesting range (e.g., $2\le n \le \sqrt p$ and $0\le k< n-1$). Rather, the formula
needs to be analzed and the random generator should be restricted to only generate
interesting values.

It would not be significantly more expensive to have every validator perform a
predetermined finite number of checks for each assertion.

What is the minimum number $m$ of validators such that the probability of spotting an
error is less than a given threshold $q$?

Instead of using random numbers, is there a way to coordinate a systematic exploration
of the iteration space? (Is this at all necessary?)

One could also check properties of the values stored by the contract.

Connections to random testing?

@@

\begin{itemize}
\item Identify a set of logical formulas that are amenable to the kind of checking
  described above.
\item Formalize the cost incurred by checking a formula.
\item Extend the contract language with assertions stating these logical formulas.
\item Implement the transformation: negation of the formula followed by transformation
  of existentials to picking a random value.

@@

  Observe that in each case there is a test to eliminate trivial cases, which avoids
  illegal invocations of \lstinline/randint/.
\item Each test function should come with a formula that gives an estimate of
  its effectiveness, i.e., how many times do we have to run the test to find a
  counterexample with a given probability $p$. Ideally, the formula returns a lower
  bound on the number of tests for any $p$.
\end{itemize}

\subsection{Protocol Design}
\label{sec:protocol-design}

>>>>

\newenvironment{ethereum}{%
  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=red!20,colframe=red!20]}{%
  \end{tcolorbox}}

@@

@@

\begin{itemize}
\item Formally define a set of logical formulas that are amenable to the kind of checking
  described above. In a first approximation, the interesting assertion formulas seem to
  be those with prenex universal quantification. 
\item Collect examples for interesting formulas.
\item Formalize the cost incurred by checking a formula.
\item Extend the contract language with assertions stating these logical formulas.
\item Implement the transformation: negation of the formula followed by transformation
  of existentials to picking a random value. For initial experiments, the target of the
  translation can be any language; later on, the target has to be a language hosted on
  the blockchain.

  Negation turns prenex universial quantification into prenex existential quantification. But it
  would be ineffective to generate a random integer and then check that it falls in the
  interesting range (e.g., $2\le n \le \sqrt p$ and $0\le k< n-1$). Rather, the formula
  needs to be analzed and the random generator should be restricted to only generate
  interesting values.

@@

  Observe that in each case there is a test to eliminate trivial cases, which avoids
  illegal invocations of \lstinline/randint/.
\item Each test function should come with a formula that gives an estimate of
  its effectiveness, i.e., how many times do we have to run the test to find a
  counterexample with a given probability $p$. Ideally, the formula returns a lower
  bound on the number of tests for any $p$.
\item What is the minimum number $m$ of validators such that the probability of
  spotting an error is less than a given threshold $q$?
\item Instead of using random numbers, is there a way to coordinate a systematic
  exploration of the iteration space? (Is this at all necessary?)
\item What are the connections to random testing? 
\item I think this scheme implements a Bloom filter, globally? We are testing whether a value is a
  member of an implicitly given set. We may accept a value even if it does not belong
  to the set, but if we reject a value, it definitely does not belong to the set.
\end{itemize}

\subsection{Blockchain-Specific Tasks}
\label{sec:blockch-spec-tasks}

\begin{ethereum}
  Ethereum-specific remarks:
  Implementing this scheme requires adding instructions to the virtual machine. One
  instruction would generate a random number (EVM uses 512 bit unsigned integer IIRC),
  the other would raise an overriding failure. This failure will have to indicate that
  validation of the contract fails.
\end{ethereum}

\begin{itemize}
\item The VM interpreter needs to be extended with these instructions.
\item The translation (of formulas) needs to retargetted to VM code. This may have to
  be hacked into the compiler.
\item The translation has to be amended so that counterexamples can also be checked.
\item (extension) Every validator could perform a predetermined finite number of checks for each
  assertion. This would not be significantly more expensive. Moreover, the number of
  checks could be set individually depending on the contract and on the size of the
  input.
\item (extension) One could also check properties of the values stored with the contract.
\end{itemize}
\subsection{Protocol Design}
\label{sec:protocol-design}

\begin{changethis}
  Tezos specific remarks: here the number of validators is really small compared to
  Bitcoin or Ethereum
  \begin{itemize}
  \item Maybe we need to propose and validate the same block multiple times before it
    gets accepted. E.g, a transaction containing this feature is only accepted if it is
    revalidated $n$ times.
  \item Extra validators only for this kind of transaction.
  \item Maybe each validator needs to conduct multiple test runs.
  \item Maybe test runs need to be coordinated
  \end{itemize}
\end{changethis}

\begin{itemize}
\item   Once a counterexample has been found,
  this counterexample is broadcast to the network where it can be verified by further
  nodes.  I think this broadcast and the reaction to this broadcast requires an update of
  the protocol: all receivers of the broadcast will check the counterexample and revert
  the transaction. 
\item A transaction using counterexamples has to
  wait for a couple of cycles in the transaction pool until the node feels safe to
  include the transaction in a block.
  The number of cycles to wait depends on the contract and on the size of the input.
  So that number should be variable.
\item (low priority) This approach poses the danger of a DOS attack. Someone could post a false
  counterexample, which does not really lead to a failure. If everyone switches to
  check this proposed counterexample instead of generating a random number, then the
  real problem may remain hidden. So, there has to be a component in the protocol that
  penalizes someone who broadcasts a false counterexample. Moreover, not all validators
  should switch to recheck a proposed counterexample; or they could check it in
  addition to their own random check.

  More precisely:
  \begin{itemize}
  \item Every validator must run at least one (round of) random tests.  
  \item If it receives a proposed counterexample, it must check the counterexample,
    too.
  \item It remains to investigate a countermeasure for nodes generating a flood of
    false counterexamples, which would keep every validator busy. Perhaps checking of
    the counterexample should also be done probabilistically.
  \end{itemize}
\end{itemize}

\section{Rejected Ideas}
\label{sec:rejected-ideas}

This section contains some ideas that we discussed but rejected. Feel free to resurrect
any of this if you find new arguments.

\begin{itemize}
\item Q: How about instead of generating and checking random number, it could
  store a Bloom filter for already checked number, so a validator can
  generate a random number and check it if it is rejected by Bloom filter,
  but this also happen that there will be some number never get checked
  because it is rejected by Bloom filter, but actually never been checked.

  A: This is not a good idea for several reasons.
  \begin{enumerate}
  \item It's not clear where we should store the Bloom filter. It has to be available
    everywhere, so it has to be stored on the blockchain.
  \item If it's on the blockchain, then it serializes the tests for counterexamples:
    the point of the scheme is to test many random values concurrently without
    interference.
  \item The Bloom filter errs in the wrong direction. It may accept elements that have
    not been added before, which means that elements that should be tested are not. We
    would need a co-Bloom filter that errs in the other direction: it should reject all
    elements that have not been added, and it may also reject elements that have been added.
  \end{enumerate}

\end{itemize}
